\section{Data}

% Look into EPIC data set (Exeter, and others; )

Data sources are still being investigated.

For the pilot work we will use the US-based database from the \emph{National Alzheimer's Coordinating Center\footnote{\url{https://naccdata.org/}}}. This will likely use just structure data (no free text) but will allow a general test of whether machine learning models may predict particular subgroup diagnoses at different teams, and test whether we can predict the \emph{next investigation} at each team.

For the follow-on funded piece of research we are investigating sources of data in the UK. These may be used stand-alone or used to fine-tune/enhance a model first trained on US data (so-called \emph{transfer learning} where machine learning models are first built on large data sets of relevant but imperfect data, and then fine-tuned using smaller data sets of more ideal data).

Potential sources of UK or international data include:

\begin{itemize}
    \item Clinical Record Interactive Search (CRIS\footnote{\url{https://www.maudsleybrc.nihr.ac.uk/facilities/clinical-record-interactive-search-cris/}}) from the Maudsley Biomedical Research Centre (BRC). This data set is a comprehensive data set including the potential to use free-text models. Access to this data is uncertain, and will need to be established.
    \item Primary data collection (e.g. from local memory clinics) - the least attractive option!
    \item Additional non-comprehensive data sets that may be used to inform the model include:
    \begin{itemize}
        \item NHS Digital Recorded Dementia Diagnoses\footnote{\url{https://digital.nhs.uk/data-and-information/publications/statistical/recorded-dementia-diagnoses/}}
        \item Office of Statistics Dementia Profile Updates\footnote{\url{https://www.gov.uk/government/statistics/dementia-profile-updates}}
        \item Kaggle data sets\footnote{\url{https://www.kaggle.com/search?q=Dementia+in\%3Adatasets}}
    \end{itemize}
\end{itemize}

Imagining data sets include:

\begin{itemize}
    \item Alzheimer’s Disease Neuroimaging Initiative (ADNI\footnote{url{https://adni.loni.usc.edu/}})
    \item Parkinson's Progression Markers Initiative (PPMI\footnote{\url{https://www.ppmi-info.org/}})
\end{itemize}
